---
title: "Extend DataRaft through its S3 protocols"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extend DataRaft through its S3 protocols}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", error = FALSE)
library(dataraft.core)
```

An adapter is an ordinary R object with S3 methods. It owns access to its external
system. Products and workflows stay independent of that system.

| Protocol | Responsibility |
|---|---|
| `dr_check_component()` | Validate configuration without publishing data |
| `dr_inspect()` | Return a safe description, without credentials or input rows |
| `dr_capabilities()` | Declare what the adapter actually supports |
| `dr_read_source()` | Return a readable table |
| `dr_execute_transform()` | Prepare a table |
| `dr_run_quality()` | Evaluate a quality rule |
| `dr_write_target()` | Persist data after checks pass |
| `dr_publish_metadata()` | Deliver metadata with explicit failure handling |

## A source

This small source keeps a table in an environment, standing in for an external
client. Inspection exposes a name, not the environment's data.
The explicit registrations make the methods available when this article executes
in a local environment. A package uses NAMESPACE S3 registrations instead.

```{r}
example_source <- function(client, name) {
  structure(list(client = client, name = name), class = "example_source")
}
dr_check_component.example_source <- function(x, ...) {
  stopifnot(is.environment(x$client), is.character(x$name), length(x$name) == 1L)
  invisible(x)
}
dr_read_source.example_source <- function(source, ...) {
  dr_check_component(source)
  source$client[[source$name]]
}
dr_inspect.example_source <- function(x, ...) list(type = "example", name = x$name)
dr_capabilities.example_source <- function(x, ...) {
  dr_component_capabilities(read = TRUE, write = FALSE, lazy = FALSE,
    transactions = FALSE, immutable = FALSE)
}
for (generic in c("dr_check_component", "dr_read_source", "dr_inspect", "dr_capabilities")) {
  registerS3method(generic, "example_source",
    get(paste0(generic, ".example_source")), envir = asNamespace("dataraft.core"))
}
client <- new.env(parent = emptyenv())
client$orders <- data.frame(id = 1:2, amount = c(10, 20))
source <- example_source(client, "orders")
dr_inspect(source)
```

## A target and its quality gate

The writer receives checked data plus descriptive context. It must throw on
failure. Its return value describes the output and must not leak credentials.
This example deliberately claims no transaction or immutability guarantees.

```{r}
example_target <- function(client, name) {
  structure(list(client = client, name = name), class = "example_target")
}
dr_check_component.example_target <- dr_check_component.example_source
dr_inspect.example_target <- dr_inspect.example_source
dr_capabilities.example_target <- function(x, ...) {
  dr_component_capabilities(read = FALSE, write = TRUE, lazy = FALSE,
    transactions = FALSE, immutable = FALSE)
}
dr_write_target.example_target <- function(target, data, context, ...) {
  target$client[[target$name]] <- data
  list(type = "example", name = target$name, rows = nrow(data))
}
for (generic in c("dr_check_component", "dr_inspect", "dr_capabilities", "dr_write_target")) {
  registerS3method(generic, "example_target",
    get(paste0(generic, ".example_target")), envir = asNamespace("dataraft.core"))
}
product <- dr_product("orders", source) |>
  dr_add_quality(~ amount >= 0) |>
  dr_set_target(example_target(client, "checked"))
result <- dr_publish(product)
stopifnot(identical(client$checked$amount, c(10, 20)))
client$orders$amount <- c(-1, 20)
failed <- dr_run(write = FALSE, stop_on_failure = FALSE, product)
stopifnot(failed$status == "blocked")
try(dr_publish(product), silent = TRUE)
stopifnot(identical(client$checked$amount, c(10, 20)))
invisible(dr_last_failure(clear = TRUE))
```

## Package an extension

Put methods in an R package importing `dataraft.core`. For each method, use
roxygen `@export` and `@importFrom dataraft.core <generic>` so R registers it even
when the package is loaded without attaching it. Test both direct protocol calls
and a complete `dr_run(write = FALSE, stop_on_failure = FALSE)` / `dr_publish()` workflow, including a blocked input
and a failed write. Close only connections the adapter itself opened.

Use the public protocols above. Names starting with `dr_internal_` are versioned
implementation interfaces for the DataRaft family and may change before 1.0;
external extensions should not import them. Private standalone helpers are
copied from core's canonical source by the repository sync script. Stateful
execution and publication logic is not copied.

`dr_execute_target()` is the advanced execution hook for backends owning their
entire transaction lifecycle. Most extensions should implement `dr_write_target()`
and let core handle preparation and quality gates. The RDS adapter in
`dataraft.adapters` is a complete filesystem example.
