---
title: "Checked deliveries with core alone"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Checked deliveries with core alone}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", error = FALSE)
library(dataraft.core)
```

A recurring delivery should have the same columns and satisfy the same rules
on every run. A product records these requirements. A recipe records preparation;
a workflow brings the two together. This example needs no storage engine or
other DataRaft component.

```{r}
contract <- dr_contract(columns = c(id = "integer", amount = "numeric"), key = "id")
product <- dr_product("orders", contract = contract) |>
  dr_add_quality(~ amount >= 0)
recipe <- dr_recipe() |> dr_step_mutate(amount = round(amount, 2))
flow <- dr_workflow() |> dr_add_product(product) |> dr_add_recipe(recipe)
result <- dr_trial(flow, data = data.frame(id = 1:3, amount = c(10, 20, 30)))
dr_collect(result)
dr_quality_report(result)
```

A failed check returns a result from `dr_trial()`. Collecting it raises an error.
Even after using a pipe, the last failed execution remains available locally.

```{r}
try(dr_trial(flow, data = data.frame(id = 1:3, amount = c(10, -20, 30))) |>
      dr_collect(), silent = TRUE)
failed <- dr_last_failure()
dr_quality_report(failed)
dr_quality_rows(failed)
invisible(dr_last_failure(clear = TRUE))
```

Use `dr_trial()` to test without configured writers, `dr_run()` to execute the
configured workflow, and `dr_publish()` to require a destination. Storage and
catalog components are optional. The metapackage's RDS example shows a durable
next step; the extension article explains how to provide your own adapter.

## Build definitions in small steps

The constructors also accept governance metadata and advanced backend settings.
Keep these optional arguments out of the first delivery definition. Start with
shape and identity, then compose the parts that change independently:

```{r}
shape <- dr_contract("orders", columns = c(id = "integer", amount = "numeric"))
positive <- dr_quality_rule(~ amount >= 0, action = "block", threshold = 0)
orders <- dr_product("orders") |>
  dr_add_contract(shape) |>
  dr_add_quality(positive)
regions <- data.frame(id = 1:2, region = c("North", "South"))
preparation <- dr_recipe() |>
  dr_step_lookup(regions, by = "id")
flow <- dr_workflow() |>
  dr_add_product(orders) |>
  dr_add_recipe(preparation) |>
  dr_add_source(data.frame(id = 1:2, amount = c(10, 20)), name = "delivery")
dr_trial(flow) |> dr_collect()
```

`dr_contract()` needs only the identity and expected columns here; supply `key`,
`grain` or governance fields when those guarantees are known. `dr_product()` needs
only identity because attachments are separate. For `dr_quality_rule()`, use
`action` and `threshold` consistently. `severity` and `max_failure` remain
compatibility arguments, and supplying both names for the same setting fails.
For `dr_step_lookup()`, start with `source` and `by`; defaults enforce a complete
many-to-one match. Set `unmatched`, `suffix`, `engine`, `name` or `table` only when
that lookup needs a different explicit policy. The full signatures remain
available so existing positional calls do not break.

Revise definitions explicitly, then replace the appropriate slot:

```{r}
revision <- dr_extract_contract(orders) |>
  dr_update_contract(version = "2", owner = "Risk")
orders <- dr_update_contract(orders, revision)
flow <- dr_update_product(flow, orders) |>
  dr_update_source(data.frame(id = 1:2, amount = c(30, 40)), name = "delivery")
dr_extract_source(flow, "delivery")
```

`dr_contract_update()` is the deprecated compatibility spelling of contract
revision. New code uses `dr_update_contract()`. A contract revision requires a
new ID or version; changing guarantees can also require explicit key, required
column or rule review. `dr_remove_contract()` clears the attached contract,
restoring automatic shape inference, while preserving added quality rules.
`dr_add_contract()` keeps its existing replacement behavior; use
`dr_update_contract()` when an existing attachment is required. A workflow's
contract is attached to its product. Primary source helpers edit sources owned
by their argument only; extract a workflow's product to edit its own sources.
