---
title: "Checked deliveries with core alone"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Checked deliveries with core alone}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", error = FALSE)
library(dataraft.core)
```

A recurring delivery should have the same columns and satisfy the same rules
on every run. A product records these requirements. A recipe records preparation;
the product owns both. This example needs no storage engine or
other DataRaft component.

```{r}
contract <- dr_contract(columns = c(id = "integer", amount = "numeric"), key = "id")
product <- dr_product("orders", contract = contract) |>
  dr_add_quality(~ amount >= 0)
recipe <- dr_recipe() |> dr_step_mutate(amount = round(amount, 2))
flow <- product |> dr_add_recipe(recipe)
result <- dr_run(write = FALSE, stop_on_failure = FALSE, flow, data = data.frame(id = 1:3, amount = c(10, 20, 30)))
dr_collect(result)
dr_quality_report(result)
```

A failed check returns a result from `dr_run(write = FALSE, stop_on_failure = FALSE)`. Collecting it raises an error.
Even after using a pipe, the last failed execution remains available locally.

```{r}
try(dr_run(write = FALSE, stop_on_failure = FALSE, flow, data = data.frame(id = 1:3, amount = c(10, -20, 30))) |>
      dr_collect(), silent = TRUE)
failed <- dr_last_failure()
dr_quality_report(failed)
dr_quality_rows(failed)
invisible(dr_last_failure(clear = TRUE))
```

Use `dr_run(write = FALSE, stop_on_failure = FALSE)` to test without configured writers, `dr_run()` to execute the
configured workflow, and `dr_publish()` to require a destination. Storage and
catalog components are optional. The metapackage's RDS example shows a durable
next step; the extension article explains how to provide your own adapter.

## Build definitions in small steps

The constructors also accept governance metadata and advanced backend settings.
Keep these optional arguments out of the first delivery definition. Start with
shape and identity, then compose the parts that change independently:

```{r}
shape <- dr_contract(
  "orders",
  columns = c(id = "integer", amount = "numeric", region = "character")
)
positive <- dr_quality_rule(~ amount >= 0, action = "block", threshold = 0)
orders <- dr_product("orders") |>
  dr_add_contract(shape) |>
  dr_add_quality(positive)
regions <- data.frame(id = 1:2, region = c("North", "South"))
preparation <- dr_recipe() |>
  dr_step_lookup(regions, by = "id")
flow <- orders |>
  dr_add_recipe(preparation) |>
  dr_set_sources(delivery = data.frame(id = 1:2, amount = c(10, 20)))
dr_run(write = FALSE, stop_on_failure = FALSE, flow) |> dr_collect()
```

The contract describes the final output, including `region` added by the lookup.
The input delivery supplies only `id` and `amount`; checks run after preparation.

`dr_contract()` exposes identity, columns, key, rules and version. Add ownership
with `dr_contract_meta()` and validation settings with `dr_contract_policy()`.
All quality engines use `action` and `threshold`. The old `severity` and
`max_failure` arguments emit lifecycle warnings and are scheduled for removal.
Lookup constraints use dm; `native` is no longer a second engine.

Revise a contract version explicitly, then replace the attachment:

```{r}
revision <- dr_contract_update(orders$contract, version = "2", owner = "Risk")
flow <- flow |> dr_add_contract(revision) |>
  dr_set_sources(delivery = data.frame(id = 1:2, amount = c(30, 40)))
dr_collect(dr_run(flow, write = FALSE))
flow$sources$delivery
```

`dr_add_contract(x, NULL)` removes the contract. `dr_set_sources(x, delivery =
NULL)` removes one input; `sources = NULL` clears them all. These operations
never read data. The former modular workflow is now just a product; there is no
separate wrapper state or execution method. `dr_trial()` is deprecated; use
`dr_run(write = FALSE)` and choose `stop_on_failure = FALSE` explicitly when
inspecting a blocked result.
