---
title: "Checked deliveries with core alone"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Checked deliveries with core alone}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", error = FALSE)
library(dataraft.core)
```

A recurring delivery should have the same columns and satisfy the same rules
on every run. A product records these requirements. A recipe records preparation;
a workflow brings the two together. This example needs no storage engine or
other DataRaft component.

```{r}
contract <- dr_contract(columns = c(id = "integer", amount = "numeric"), key = "id")
product <- dr_product("orders", contract = contract) |>
  dr_add_quality(~ amount >= 0)
recipe <- dr_recipe() |> dr_step_mutate(amount = round(amount, 2))
flow <- dr_workflow() |> dr_add_product(product) |> dr_add_recipe(recipe)
result <- dr_trial(flow, data = data.frame(id = 1:3, amount = c(10, 20, 30)))
dr_collect(result)
dr_quality_report(result)
```

A failed check returns a result from `dr_trial()`. Collecting it raises an error.
Even after using a pipe, the last failed execution remains available locally.

```{r}
try(dr_trial(flow, data = data.frame(id = 1:3, amount = c(10, -20, 30))) |>
      dr_collect(), silent = TRUE)
failed <- dr_last_failure()
dr_quality_report(failed)
dr_quality_rows(failed)
invisible(dr_last_failure(clear = TRUE))
```

Use `dr_trial()` to test without configured writers, `dr_run()` to execute the
configured workflow, and `dr_publish()` to require a destination. Storage and
catalog components are optional. The metapackage's RDS example shows a durable
next step; the extension article explains how to provide your own adapter.
